\documentclass{letter}
\usepackage[a4paper, total={6in, 8in}]{geometry}
\signature{Grzegorz Koperwas}
\address{5 Żniwna Street \\ 42-675 Ziemięcice \\ Poland}

\longindentation=0pt
\begin{document}
\begin{letter}{CEO of \textbf{blindr}\\ 21 Sun Road \\ Los Angeles\\ California,
United States of America}

\opening{Dear Mr. Eric Wang}

I am writing to express my interest in your job offer at \textbf{blindr}. I have come across your Company while browsing the lists of the most innovative startups in Silicon Valley while having the problem of sun reflecting of my computer screen. Because of that, I was inspired to join your mission to construct the best window blinds humanly possible.

I am currently working as a fullstack developer working with python for 2 years. While I might not be the best with data analysis, my strength is creating great interfaces for end users. I think we together can create the best, the smoothest window blind user experience possible, with the right amount of usability and exclusivity they deserve.

Previously I held junior developer position creating innovative interfaces for home automation systems. During my internship I had developed Augmented Reality interfaces for various appliances, allowing users to control their home intuitively with their phones. I think such a solution might mesh well with your existing machine learning and blockchain solutions.

Currently, I want to develop my skillset to fit a more managerial role, as in my current position I find joy in developing agile work organization techniques. I like cutting the red tape in any process, so my team can deliver their solutions to the market first.

I would like to join your product development team, where I would like to apply my skillset to improve your products. You can contact me at my personal email address at admin@grzegorzkoperwas.site.


Thank you for your consideration

\closing{Sincerely,}
\end{letter}
\end{document}
